---
title: "Data cleaning"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup

```{r, warning=FALSE, message=FALSE}
# Load packages
library(foreign)
library(tidyverse)
library(here)

# Load custom functions
script_files <- list.files(here::here("R"), pattern = "\\.R$", full.names = TRUE)
purrr::walk(script_files, source)
```

# Load data

```{r, message=FALSE}
source <- read_csv(here::here("data/source/source_2024-03-30.csv"))
```

# Removing Qualtrics labels

```{r}
label <-
  source |> 
  slice(1:2)

source <- 
  source |> 
  slice(-(1:2))
```

# Exploring drop-outs

We are excluding participants who did not finished the survey. We will not share the data of these participants as promised by the consent informed form.

Define completed participants.

```{r}
dropouts <- source |> 
  dplyr::filter(!(Progress > 97 & consent == "Yes"))
```

Remove metadata columns (adjust as needed). Assume your question columns start at e.g. column 10 (change accordingly).

```{r}
question_cols <- dropouts |> 
  dplyr::select(10:ncol(dropouts))
```

Apply function to each dropout.

```{r}
dropout_points <- apply(question_cols, 1, find_dropout_point)
```

Get frequencies.

```{r}
dropout_freq <- table(factor(dropout_points, levels = colnames(question_cols)))
print(dropout_freq)
```

# Exclusions 

We exclude everyone who completed less than 98% of the survey.
Participants who completed 98% answered all questions, but did not click the submit button.

```{r}
source <-
  source |> 
  dplyr::mutate(
    Progress = as.numeric(as.character(Progress))
  )

source_filtered <- 
  source |> 
  dplyr::filter(Progress > 97 & consent == "Yes")
```

# Select needed columns

We drop the columns that contain free text since it is possible that the respondents are identifiable by the comments.

```{r}
processed <-
  source_filtered |> 
  dplyr::select(c(4:26,28,30:32,34,36:38,40,42:45,47,49:53,55,57:59,61:63,66))
```

# Save processed data

```{r}
write_csv(processed, here::here("data/processed/PreregistrationDeviation_processed_data.csv"))

label |> 
  dplyr::select(c(4:26,28,30:32,34,36:38,40,42:45,47,49:53,55,57:59,61:63,66)) |> 
  write_csv(here::here("data/processed/PreregistrationDeviation_processed_labels.csv"))
```

